### 2012-02-23 doc/Manuscript/paper.tex
Added new version of the manuscript, in LNCS format.

\documentclass{llncs}

\usepackage{proof}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{skak} % for the \qside
\usepackage{mathpartir}
\usepackage{array}
\usepackage{url}

\newcommand{\fun}[1]{\mbox{\textbf{#1}}}
\newcommand{\lb}[1]{#1_{\downarrow}}
\newcommand{\ub}[1]{#1_{\uparrow}}
\newcommand{\varset}[1]{\mbox{$\cal{#1}$}}

\sloppy

\begin{document}

\title {Range Analysis of Whole Programs}

\author{Victor Sperle Campos, Igor Rafael, Douglas do Couto Teixeira,
Fernando Magno Quint\~{a}o Pereira}

\institute{UFMG -- 6627 Ant\^{o}nio Carlos Av, 31.270-010, Belo Horizonte, Brazil
\email{\{victorsc,igor.rafael,douglas,fernando\}@dcc.ufmg.br}
}

\maketitle

%mmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmm
\begin{abstract}
This paper describes an inter-procedural range analysis algorithm that scales
up to programs with millions of assembly instructions.
Contrary to many previous techniques, we handle comparisons between variables
without resorting to expensive algorithms.
We achieve path sensitiveness by using Bodik's Extended Static Single
Assignment form as the intermediate representation.
We show that by processing the strongly connected components that constitute
the graph of dependences between variables in topological order we not only
gain in time, but also in precision.
We have implemented our technique into LLVM, an industrial strenght compiler,
and have been able to process over 4 million assembly instructions in a
few seconds.
\end{abstract}

\section{Introduction}
\label{sec:intro}

% The context: why range analysis is necessary.
The analysis of integer variables on the interval lattice has been the
canonical example of abstract interpretation since its introduction by
Cousot and Cousot~\cite{Cousot77}.
Compilers use range analysis to infer the possible values that discrete
variables may assume during the execution of programs.
This analysis has many uses.
For instance, it allows the optimizing compiler to remove from the program text
redundant overflow tests~\cite{Sol11} and unnecessary array bound
checks~\cite{Bodik00}.
Additionally, range analysis is essential not only to the bitwidth aware
register allocator~\cite{Barik06,Tallam03}, but also to more traditional
allocators that handle registers of different
sizes~\cite{Kong98,Pereira08,Scholz02}.
Finally, range analysis has also seen use in the static prediction of
branches~\cite{Patterson95}, to detect buffer overflow
vulnerabilities~\cite{Simon08,Wagner00}, to find the trip count of
loops~\cite{Lokuciejewski09}
and even in the synthesis of hardware~\cite{Cong05,Mahlke01}.

% The problem: few implementations, not extensive experimental reports.
Given this great importance, it comes as no surprise that the compiler
literature is rich in works describing in details algorithmic variations of
range analyses~\cite{Gawlitza09,Mahlke01,Stephenson00,Su05}.
On the other hand, none of these authors provide experimental evidence that
their approaches are able to deal with very large programs.
There are researchers who have implemented range analyses that scale up to
large programs~\cite{Blanchet03,Venet04}; nevertheless, because the
algorithm itself is not the main focus of their works, they neither give
details about their design choices nor provide experimental data about.
This scenario was recently changed by Oh {\em et al.}~\cite{Oh12}, who
introduced an abstract interpretation framework which handles programs with
hundreds of thousands of lines of code.
Nevertheless, Oh {\em et al.} have designed a very simple range analysis,
which does not handle comparisons between variables, for instance.
They also do not discuss the precision of their implementation, but only its
runtime.

% Our contribution.
In this paper we provide a complete description of a range analysis algorithm,
and show extensive experimental data that justifies our engineering choices.
Our main algorithmic contribution on top of previous works is a three-phases
approach to handle comparisons between variables without resorting to any
exponential time technique.
The few publicly available implementations of range analyses that we are
aware of, such as those in Java Hotspot~\footnote{} and gcc~\footnote{} only
deal with comparisons between variables and constants.
Even the theoretical works, such as Su and Wagner's~\cite{Su05} or Gawlitza
{\em et al.}'s~\cite{Gawlitza09} suffer from this limitation.
This deficiency is one of the reasons explaining why none of these works has
made their way into industrial-strength compilers.
Two other insights allow our implementation to scale up to very large
programs.
Firstly, we use Bodik's Extended Static Single Assignment (e-SSA) form to
perform range analysis sparsely.
This program representation ensures that the interval associated with
a variable is constant along its entire live range, and allows us to compute
path-sensitive results.
Secondly, we process the strongly connected components that underlines our
constraint system in topological ordering.
It is well-know that this technique is essential to speedup constraint solving
algorithms~\cite[Sec 6.3]{Nielson99}; however, we show that a careful
propagation of information along strong components not only gives us speed, but
also improves the precision of our results.

% The history of our contribution.
This work concludes a two years long effort to produce a solid and scalable
implementation of range analysis.
Our first endeavor to implement such an algorithm was based on Su and
Wagner's constraint system, which can be solved exactly in polynomial
time~\cite{Su04,Su05}.
However, although we could use their formula to handle a subset of C-like
constructs, their description of how to deal with loops was
``not very explicit"~\cite[p.422]{Gawlitza09}.
Thus, in order to solve loops we adopted Gawlitza
{\em et al.}'s~\cite{Gawlitza09} approach.
This technique uses the Bellman-Ford algorithm to detect increasing or
decreasing cycles in the constraint system, and then saturates these cycles
via a simple widening operator.
A detailed description of our implementation has been published by
Couto and Pereira~\cite{Couto11}.
Nevertheless, the inability to handle comparisons between variables, and the
cubic complexity of the Bellman-Ford method eventually leads us to seek
alternative solutions to range analysis.
This quest reached a pinnacle in the present work.

% Our experimental numbers.
We have implemented our algorithm in the LLVM compiler~\cite{Lattner04}, and
have used it to analyze a test suite with 1.4 million lines of code.
In Section~\ref{sec:exp} we provide empirical evidence showing that our
implementation is fast: it analyzes the gcc source code in XX seconds, for
instance.
It is also precise: our results are similar to Stephenson {\em et al}'s,
even though, contrary to them, we do not assume that the program is correct.
Furthermore, we have been able to find tight bounds to many of the examples
used by Costan {\em et al.}~\cite{Costan05} and Lakhdar
{\em et al.}~\cite{Lakhdar11}, even though they use much more costly methods.

% The rest of this paper.

\section{Background}
\label{sec:bck}

% Define the lattice, the constraints and the valuation I.
Following Gawlitza {\em et al.}'s notation, we shall be performing arithmetic
operations over the complete lattice
$\cal{Z} = \mathbb{Z} \cup \{-\infty, +\infty\}$, where the ordering is
naturally given by $-\infty < \ldots < -2 < -1 < 0 < 1 < 2 < \ldots +\infty$.
For any $x > -\infty$ we define:

\begin{tabular}{lcl}
$x + \infty = \infty, x \neq -\infty$ & \mbox{\hspace{0.1cm}} & $x - \infty = - \infty, x \neq +\infty$ \\
$x \times \infty = \infty$ if $x > 0$ & & $x \times \infty = -\infty$ if $x < 0$ \\
$0 \times \infty = 0$ & & $(-\infty) \times \infty = \ $ not defined $$ \\
\end{tabular}

From the lattice $\varset{Z}$ we define the product lattice
$\varset{Z}^2$, which is partially ordered by the subset relation
$\sqsubseteq$.
$\varset{Z}^2$ is defined as follows:

\begin{equation*}
\varset{Z}^2 = \emptyset \cup \{[z_1, z_2] | \ z_1,z_2 \in \varset{Z},
\ z_1 \leq z_2, \  -\infty < z_2 \}
\end{equation*}

Given an interval $\iota = [l, u]$, we let $\lb{\iota} = l$, and
$\ub{\iota} = u$.
We let \varset{V} be a set of variables, and
$I: \varset{V} \mapsto \varset{Z}^2$ a
mapping from variables to intervals in $\varset{Z}^2$.
Our objective is to solve a constraint system $C$, formed by instances of the
six different types of constraints seen in Figure~\ref{fig:eval_function}
(left).
Figure~\ref{fig:eval_function} (right) defines a valuation function $e$ that
computes $Y = f(\ldots)$, given I.
Armed with these concepts, we define the range analysis problem as follows:

\begin{definition}
\label{def:rcp}
\textsc{Range Analysis Problem} \\
\textbf{Input:} a set $C$ of constraints ranging over a set \varset{V} of
variables. \\
\textbf{Output:} a mapping I such that, for any variable
$V \in \varset{V}$, e(V) = I[V].
\end{definition}

\begin{figure}[t!]
\begin{small}
\begin{eqnarray*}
\begin{array}{r@{\hspace{0.5cm}}c}
Y = [l, u]
&
e(Y) = [l, u]
\\
\\
Y = \phi (X_1, X_2)
&
\inferrule{I[X_1]=[l_1, u_1] \\ I[X_2]=[l_2, u_2]}
{e(Y) = [\mbox{min}(l_1, l_2), \mbox{max}(u_1, u_2)]}
\\
\\
Y = X_1 + X_2
&
\inferrule{I[X_1]=[l_1, u_1] \\ I[X_2]=[l_2, u_2]}
{e(Y) = [l_1 + l_2, u_1 + u_2]}
\\
\\
Y = X_1 \times X_2
&
\inferrule{I[X_1]=[l_1, u_1] \\ I[X_2]=[l_2, u_2] \\ L = \{l_1l_2, l_1u_2, u_1l_2, u_1u_2\}}
{e(Y) = [\mbox{min}(L), \mbox{max}(L)]}
\\
\\
Y = aX + b
&
\inferrule{I[X]=[l, u] \\ k_l = al + b \\ k_u = au + b}
{e(Y) = [\mbox{min}(k_l, k_u), \mbox{max}(k_l, k_u)]}
\\
\\
Y = X \sqcap [l', u']
&
\inferrule{I[X]=[l, u]}
{e(Y) \leftarrow [\mbox{max}(l, l'), \mbox{min}(u, u')]}
\end{array}
\end{eqnarray*}
\caption{\label{fig:eval_function}
The constraint system that produces instances of the range analysis problem.}
\end{small}
\end{figure}

We will use the program in Figure~\ref{fig:ex1}(a) as the running example
to illustrate our range analysis.
Figure~\ref{fig:ex1}(b) shows the same program in SSA form, and
Figure~\ref{fig:ex1}(c) outlines the constraints that we extract from this
program.
There is a clear correspondence between instructions and constraints.
A possible solution to the range analysis problem is given in
Figure~\ref{fig:ex1}(d).
This is a very conservative solution.
As we will see shortly, we can improve this solution substantially by using
a more sophisticated program representation.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=0.7\textwidth]{images/ex1}
\end{center}
\caption{\label{fig:ex1}
(a) Example program.
(b) Control Flow Graph (CFG) in SSA form.
(c) Constraints that we extract from the program.
(d) A possible solution to the range analysis problem.}
\end{figure}

\section{Range Analysis in Three Steps}
\label{sec:algo}

% Program representation
% Extracting constraints
% Building the Constraint graph (pseudo-edges)
% Macro algorithm
% Micro algorithm

In this section we explain the algorithm that we use to solve the range
analysis problem for an entire program.
This algorithm involves a number of steps.
First, we convert the program to a suitable intermediate representation that
makes it easier to extract constraint.
From these constraints, we build a dependence graph that allows us to do
range analysis sparsely.
Finally, we solve the constraints applying different fix-point iterators on
this dependence graph.
Figure~\ref{fig:algorithm} gives a global view of this algorithm.
Some of the steps in the algorithm are optional.
They improve the precision of the range analysis, at the expenses of a longer
running time.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=0.8\textwidth]{images/algorithm}
\end{center}
\caption{\label{fig:algorithm}
Our algorithm to solve the range analysis problem.
Rounded boxes are optional steps.}
\end{figure}

\subsection{Choosing a Program Representation}
\label{sub:rep}

The solution that we found to the range analysis problem in Figure~\ref{fig:ex1}
is imprecise because we did not take conditional tests into considerations.
Branches gives us information about the ranges that some variables assume, but
only at {\em specific} program points.
For instance, given a test such as $(k_1 < 100)?$ in  Figure~\ref{fig:ex1}(b),
we know that $I[k_1] \sqsubseteq [-\infty, 99]$ whenever the condition is true.
In order to encode this information, we might split the live range of $k_1$
right after the branching point; thus, creating two new variables, one at the
path where the condition is true, and another where it is false.
There is a program representation, introduced by Bodik
{\em et al.}~\cite{Bodik00}, that performs this live range splitting:
the {\em Extended Static Single Assignment} form, or e-SSA for short.
In this paper, we have experimented with two different flavors of e-SSA form,
which we call {\em standard} and {\em non-pruned}.

We will show first how we convert a program into standard e-SSA form.
Let $(v < c)?$ be a conditional test, and let $l_t$ and $l_f$ be labels in
the program, such that $l_t$ is the target of the test if the condition is true,
and let $l_f$ is the target when the condition is false.
If $l_t$ dominates any use of $v$, then we insert at $l_f$ a copy
$v_f = v \sqcap [-\infty, c-1]$, where $v_f$ is a fresh name.
We then rename every use of $v$ that is dominated by $l_f$ to $v_f$.
Dually, if $l_t$ dominates any use of $v$, then we insert at $l_t$ a copy
$v_t = v \sqcap [c, +\infty]$, and rename variables accordingly.
If the test uses two variables, e.g., $(v_1 < v_2)?$, then we create
intersections bound to {\em futures}.
We insert, at $l_f$, $v_{1f} = v_1 \sqcap [\fun{ft}(v_2), +\infty]$,
and $v_{2f} = v_2 \sqcap [-\infty, \fun{ft}(v_1) - 1]$.
Similarly, at $l_v$ we insert
$v_{1v} = v_1 \sqcap [-\infty, \fun{ft}(v_2) - 1]$
and $v_{2f} = v_2 \sqcap [\fun{ft}(v_1), +\infty]$.
We use the notation $\fun{ft}(v)$ to denote the {\em future} bounds of a
variable.
As we will show in Section~\ref{sub:micro}, once the growth pattern of $v$ is
known, we can replace $\fun{ft}(v)$ by an actual value.
Once we are done placing copies, we insert $\phi$-functions into the
transformed program to convert it to SSA form.
This last step prevents that two different names given to the same original
variable be simultaneously alive at the program code.
Figure~\ref{fig:ex_standard_eSSA} shows our running example changed into
standard e-SSA form.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=0.5\textwidth]{images/ex_standard_eSSA}
\end{center}
\caption{\label{fig:ex_standard_eSSA}
The control flow graph from Figure~\ref{fig:ex1}(b) converted to standard
e-SSA form.}
\end{figure}

The standard e-SSA form serves well analysis that need range information
at the use site of variables, such as conditional constant propagation,
redundant code elimination and detection of buffer overflows.
On the other hand, there exist compilation algorithms that require range
information at the whole live range of variables.
Examples of such algorithms include the family of
bitwidth aware register allocators~\cite{Barik06,Tallam03,Pereira08} and
hardware synthesizers~\cite{Cong05,Mahlke01,Stephenson00}.
In order to provide extra precision to these algorithms, we may work with a
non-pruned version of e-SSA form, which we produce by simply inserting copies
after every conditional test, and then reconverting the entire program to
SSA form.
Figure~\ref{fig:ex_non_pruned} illustrates the difference between these two
program representations.
Notice that the non-pruned flavor might introduce dead variables in the
program code, which can be removed by standard dead code elimination.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=0.9\textwidth]{images/ex_non_pruned}
\end{center}
\caption{\label{fig:ex_non_pruned}
(a) Example program.
(b) Standard e-SSA form.
(c) Non-pruned e-SSA form.}
\end{figure}

\subsection{Extracting Constraints}
\label{sub:constraints}

The constraints shown in Figure~\ref{fig:eval_function} let us handle
many different assembly instructions.
We use $\phi$ assignments, such as $Y = \phi(X_1, X_2)$ to represent
the $\phi$-functions typically found in SSA form programs.
As we have discussed in Section~\ref{sub:rep}, we use intersection
constraints to bound the variables created due to live range splitting
at conditionals.
The widening analysis that we will introduce in Section~\ref{sub:micro}
require monotonic transfer functions.
Many assembly operations, such as modulus or division, do not afford us
this property.
However, several of these non-monotonic instructions have a conservative
translation to our constraint system, as we show in
Figure~\ref{fig:constraints}.

\begin{figure}[t!]\begin{center}
\begin{tabular}{|l@{\hspace{0.2cm}}l@{\hspace{0.3cm}}l|} \hline
\textbf{Description} & \textbf{Operation} & \textbf{Constraint} \\ \hline
Constant & $v = c$ & $V = [c, c]$ \\
Assignment & $v_1 = v_2$ & $V_1 = V_2$ \\
Multiplication & $v_1 = v_2 * c$ & $V_1 = c V_2$ \\
Int. division & $v_1 = v_2 \ / \ v_3$ & $V_1 = V_2$ \\
 & & $V_1 = - V_2$ \\
Modulus & $v_1 = v_2 \ \% \ c$ & $V_1 = [0, c - 1]$ \\
Bitwise and & $v_1 = v_2 \ \& \ c$ & $V_1 = v_2 \sqcap [0, c]$ \\
Bitwise or & $v_1 = v_2 \ | \ v_3$ & $V_1 = V_2 + V_3$ \\
Left shift & $v_1 = v_2 \ \qside \ c$ & $V_1 = 2^c V_2$ \\ \hline
\end{tabular}
\end{center}
\caption{\label{fig:constraints}Example of constraint derivation rules.
We let $c$ denote a positive constant, $\{v, v_1, v_2\}$ are program
variables and $\{V, V_1, V_2\} \subset \cal{V}$.}
\end{figure}

\subsection{The Dependence Graph}
\label{sub:graph}

The main data structure that we use to solve the range analysis problem is
a variation of Ferrante {\em et al.}'s {\em program dependence
graph}~\cite{Ferrante87}.
For each constraint variable $V$ we create a variable node $N_v$.
For each constraint $C$ we create a constraint node $N_c$.
We add an edge from $N_v$ to $N_c$ if the name $V$ is used in $C$.
We add an edge from $N_c$ to $V$ if the constraint $C$ defines the name
$V$.
Figure~\ref{fig:ex_graph} shows the dependence graph that we build for the
e-SSA form program given in Figure~\ref{fig:ex_standard_eSSA}.
If $V$ is used by $C$ as the input of a future, then the edge from
$N_v$ to $N_c$ represents a {\em control
dependence}~\cite[p.323]{Ferrante87}.
All the other edges denote {\em data dependences}~\cite[p.322]{Ferrante87}.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=0.7\textwidth]{images/ex_graph}
\end{center}
\caption{\label{fig:ex_graph}
The dependence graph that we build to the program in
Figure~\ref{fig:ex_standard_eSSA}.}
\end{figure}

\subsection{Finding Ranges: the Macro Algorithm}
\label{sub:macro}

Having built the dependence graph, we proceed to solve the range analysis
problem by finding all the strongly connected components (SCCs) of the
dependence graph, and reducing this graph to a directed acyclic graph.
We perform this last step by collapsing every SCC into a single node.
We then sort the resulting DAG topologically, and apply the analyses from
Section~\ref{sub:micro} on every SSC in topological order.
Once we solve the range analysis problem for a SCC, we propagate the
intervals that we found to the variable nodes at the {\em frontier} of this
SCC.
A variable node $N_v$ is said to be in the frontier of a strongly connected
component $S$ if:
(i) $N_v \notin S$; and
(ii) there exists a variable node $V' \in S$, and a constraint node $N_c$,
such that $V' \leftarrow N_c$, and $N_c \leftarrow V$.
This propagation gives us a very modular algorithm:
when analyzing a strongly connected component $S$ we can rest assured that
any influence that $S$ might suffer from nodes outside it has been already
taken into consideration.

\subsection{Finding Ranges: the Micro Algorithm}
\label{sub:micro}

Given a strongly connected component of the dependence graph with $N$ nodes,
we solve the range analysis problem in three-steps:
\begin{enumerate}
\item Run widening analysis: $O(N)$.

\item Fix intersections: $O(N)$.

\item Run narrowing analysis: $O(N^2)$.

\end{enumerate}
However, before we start, we remove the control dependence edges from the
strongly connected component, as they have no semantics to our transfer
functions.

\paragraph{Widening Analysis.}

The first step of our algorithm consists in determining the growth behavior of
the interval bound to each constraint variable.
We accomplish this task via Cousot and Cousot's widening
operator~\cite[p.247]{Cousot77}.
The possible behaviors of an interval are:
(i) does not change;
(ii) grows towards $+\infty$;
(iii) grows towards $-\infty$; and
(iv) grows in both directions.
The lattice of this dataflow analysis, plus its meet operator is given in
Figure~\ref{fig:growth_analysis}.
Because the lattice has height three, the intervals bound to each variable can
change at most three times.

\begin{figure}[t!]
\begin{center}
\begin{tabular}{c@{\hspace{1.5cm}}c}
\begin{minipage}{2.4cm}
\includegraphics{images/growth_lattice}
\end{minipage}
&
\begin{minipage}{6cm}
\begin{small}
\begin{eqnarray*}
\begin{array}{c}
\inferrule{I[Y] = [\bot, \bot]}{I[Y] \leftarrow e(Y)}
\\
\\
\inferrule{\lb{e(Y)} < \lb{I[Y]} \\ \ub{e(Y)} > \ub{I[Y]}}
{I[Y] \leftarrow [-\infty, +\infty ]}
\\
\\
\inferrule{\lb{e(Y)} < \lb{I[Y]}}
{I[Y] \leftarrow [-\infty, \ub{I[Y]}]}
\\
\\
\inferrule{\ub{e(Y)} > \ub{I[Y]}}
{I[Y] \leftarrow [\lb{I[Y]}, +\infty]}
\end{array}
\end{eqnarray*}
\end{small}
\end{minipage}
\end{tabular}
\end{center}
\caption{\label{fig:growth_analysis}
(Left) The lattice of the widening analysis.
(Right) Cousot and Cousot's widening operator. We evaluate the rules from top 
to bottom, and stop upon finding a pattern matching.}
\end{figure}

\paragraph{Fixing futures.}

The ranges found by the widening analysis tells us which variables have fixed
bounds, independent on the intersections in the constraint system.
Thus, we can use actual limits to replace intersections bounded by futures.
Figure~\ref{fig:fix_intersects} shows the rules to perform these substitutions.
In order to correctly replace a future $\fun{ft}(V)$ that limits a variable
$V'$, we need to have already applied the widening analysis onto $V$.
Had we considered only data dependence edges, then it would be possible
that $V$ be analyzed before $V'$.
However, because of control dependence edges, this case cannot happen.
The control dependence edges ensure that any topological ordering of the
constraint graph either places $N_v$ before $N_{v'}$, or places these nodes
in the same strongly connected component.
For instance, in Figure~\ref{fig:ex_graph}, variables $J_1$ and $I_t$ are in
the same SCC only because of the control dependence edges.

\begin{figure}[t!]
\begin{center}
\begin{eqnarray*}
\begin{array}{c}
\inferrule{Y = X \sqcap [l, \fun{ft}(V) + c] \\ \ub{I[V]} = u}
{Y = X \sqcap [l, u + c]} \mbox{\hspace{0.3cm}} u, c \in \mathbb{Z} \cup \{-\infty, +\infty\}
\\
\\
\inferrule{Y = X \sqcap [\fun{ft}(V) + c, u] \\ \lb{I[V]} = l}
{Y = X \sqcap [l + c, u]} \mbox{\hspace{0.3cm}} l, c \in \mathbb{Z} \cup \{-\infty, +\infty\}
\end{array}
\end{eqnarray*}
\end{center}
\caption{\label{fig:fix_intersects}Rules to replace futures by actual
bounds. $S$ is the interval bound to each variable after the widening
analysis.}
\end{figure}

%\begin{figure}[t!]
%\begin{equation*}
%S[X] \leftarrow
%\begin{cases}
%``\ ? \ " \ \ \mbox{if} \ I[X] = [-\infty, +\infty]\\
%``\downarrow" \ \mbox{if} \ I[X] = [-\infty, c], c \in \mathbb{Z} \\
%``\uparrow" \ \mbox{if} \ I[X] = [c, +\infty],  c \in \mathbb{Z} \\
%`` \ 0 \ " \ \ \mbox{if} \ I[X] = [c_1, c_2], c_1, c_2 \in \mathbb{Z}
%\end{cases}
%\end{equation*}
%\caption{\label{fig:st_mem}The growth state of each variable.}
%\end{figure}

\paragraph{Narrowing Analysis.}

The widening analysis associates very conservative bounds to each variable.
Thus, the last step of our algorithm consists in narrowing these intervals.
We accomplish this step via Cousot and Cousot's narrowing
operator~\cite[248]{Cousot77}, which we show in
Figure~\ref{fig:crop_analysis}.

\begin{figure}[t!]
\begin{center}
\begin{eqnarray*}
\begin{array}{c}
\inferrule{\lb{I[Y]} = -\infty \\ \lb{e(Y)} > -\infty}
{I[Y] \leftarrow [\lb{e(Y)}, \ub{I[Y]}]}
\\
\\
\inferrule{\lb{I[Y]} > \lb{e(Y)}}
{I[Y] \leftarrow [\lb{e(Y)}, \ub{I[Y]}]}
\\
\\
\inferrule{\ub{I[Y]} = +\infty \\ \ub{e(Y)} < +\infty}
{I[Y] \leftarrow [\lb{I[Y]}, \ub{e(Y)}]}
\\
\\
\inferrule{\ub{I[Y]} < \ub{e(Y)}}
{I[Y] \leftarrow [\lb{I[Y]}, \ub{e(Y)}]}
\end{array}
\end{eqnarray*}
\end{center}
\caption{\label{fig:crop_analysis}Cousot and Cousot's narrowing operator.}
\end{figure}

\paragraph{Example}

Continuing with our example, Figure~\ref{fig:ex_partition_grow_crop} shows
the application of our algorithm on the last strongly connected component of
Figure~\ref{fig:ex_graph}.
Notice that upon meeting this SCC, we have already found the interval
$[0, 0]$ to $I_0$ and the interval $[100, 100]$ to $J_0$, as we show in
Figure~\ref{fig:ex_graph}(a).
We are not guaranteed to find the least fix point of a constraint system.
However, in this example we did it.
We emphasize that finding this tight solution was only possible because of
the topological ordering of the constraint graph in
Figure~\ref{fig:ex_graph}.
Had we applied the widening operator onto the whole graph, then we would
have found out that variable $J_0$ is bound to $[-\infty, +\infty]$,
because
(i) it receives its interval directly from variable $K_t$, which is upper
bounded by $+\infty$, and
(ii) it is part of a negative cycle.
On the other hand, because we only analyze the $J$'s SCC after we have
analyzed $K$'s, $K$ only contribute the constant range $[0, 99]$ to $J_0$.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=0.8\textwidth]{images/ex_partition_grow_crop}
\end{center}
\caption{\label{fig:ex_partition_grow_crop}
Four snapshots of the last SCC of Figure~\ref{fig:ex_partition_grow_crop}.
(a) After removing control dependence edges, immediately before applying
Cousot and Cousot's operators.
(b) After running the widening analysis.
(c) After fixing the intersections.
(d) After running the narrowing analysis.}
\end{figure}

%\subsection{Alternative Narrowing Operators}
%\label{sub:alt}
%
%Cousot and Cousot's narrowing operator, given in
%Figure~\ref{fig:crop_analysis} is not guaranteed to find the least fixpoint
%of a system of constraints.
%In order to improve the bounds, trading time for precision, we have
%experimented with two other operators.
%The first, is the evaluation function itself, which is given in
%Figure~\ref{fig:eval_function}.
%By iterating it until reaching a fixpoint, we are guaranteed to find the
%optimal solution to the narrowing analysis.
%However, this approach is not practical, because it is equivalent to
%concretely interpreting the program.
%The second alternative that we have tested consists in applying the
%evaluation function on each constraint at most once per intersection.
%We call this approach {\em depth-first narrowing}, because, starting from
%an intersection we visit every node that might be influenced by it.
%After removing control dependence edges, if we are left with a strongly
%connected component containing only {\em ascencing constraints}, that
%is, constraints that only increase the upper bounds of intervals, then this
%approach is guaranteed to find the least fixpoint.
%The same is true if we have a SCC with only {\em descending constraints}.

\bibliographystyle{plain}
\bibliography{bibtex/references}

\end{document}