### 2012-03-22 doc/Manuscript/paper.tex
Added small changes to the paper, to keep it under 15 pages.

% Explain that we get the topological ordering for free via nuutila's algorithm.
\documentclass{llncs}

\usepackage{proof}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{skak} % for the \qside
\usepackage{mathpartir}
\usepackage{array}
\usepackage{url}

\newcommand{\fun}[1]{\mbox{\textbf{#1}}}
\newcommand{\lb}[1]{#1_{\downarrow}}
\newcommand{\ub}[1]{#1_{\uparrow}}
\newcommand{\varset}[1]{\mbox{$\cal{#1}$}}

\sloppy

\begin{document}

\title {Range Analysis of Whole Programs}

\author{Victor Sperle Campos, Igor Rafael de Assis Costa, Douglas do Couto
Teixeira, Fernando Magno Quint\~{a}o Pereira}

\institute{UFMG -- 6627 Ant\^{o}nio Carlos Av, 31.270-010, Belo Horizonte, Brazil
\email{\{victorsc,igor.rafael,douglas,fernando\}@dcc.ufmg.br}
}

\maketitle

%mmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmm
\begin{abstract}
This paper describes an inter-procedural range analysis algorithm that scales
up to programs with millions of assembly instructions.
Contrary to many previous techniques, we handle comparisons between variables
without resorting to relational lattices or expensive algorithms.
We achieve path sensitiveness by using Bodik's Extended Static Single
Assignment form as the intermediate representation.
%We show that by processing the strongly connected components that constitute
%the graph of dependences between variables in topological order we not only
%gain in time, but also in precision.
We claim to present in this paper the most extensive empirical evaluation of
range analysis in an industrial strength compiler.
We have implemented our technique in LLVM and have been able to process
programs totaling 2.72 million lines of C in a few seconds.
We do global program analysis, and achieve a simple form of context
sensitiveness via function inlining.
\end{abstract}

\section{Introduction}
\label{sec:intro}

% The context: why range analysis is necessary.
The analysis of integer variables on the interval lattice has been the
canonical example of abstract interpretation since its introduction in
Cousot and Cousot's seminal paper~\cite{Cousot77}.
Compilers use range analysis to infer the possible values that discrete
variables may assume during program execution.
This analysis has many uses.
For instance, it allows the optimizing compiler to remove from the program text
redundant overflow tests~\cite{Sol11} and unnecessary array bound
checks~\cite{Bodik00}.
Additionally, range analysis is essential not only to the bitwidth aware
register allocator~\cite{Barik06,Tallam03}, but also to more traditional
allocators that handle registers of different
sizes~\cite{Kong98,Pereira08,Scholz02}.
Finally, range analysis has also seen use in the static prediction of
branches~\cite{Patterson95}, to detect buffer overflow
vulnerabilities~\cite{Simon08,Wagner00}, to find the trip count of
loops~\cite{Lokuciejewski09}
and even in the synthesis of hardware~\cite{Cong05,Mahlke01}.

% The problem: few implementations, not extensive experimental reports.
Given this great importance, it comes as no surprise that the compiler
literature is rich in works describing in details algorithmic variations of
range analyses~\cite{Gawlitza09,Mahlke01,Stephenson00,Su05}.
On the other hand, none of these authors provide experimental evidence that
their approaches are able to deal with very large programs.
There are researchers who have implemented range analyses that scale up to
large programs~\cite{Blanchet03,Venet04}; nevertheless, because the
algorithm itself is not the main focus of their works, they neither give
details about their design choices nor provide experimental data about it.
This scenario was recently changed by Oh {\em et al.}~\cite{Oh12}, who
introduced an abstract interpretation framework which processes programs with
hundreds of thousands of lines of code.
Nevertheless, Oh {\em et al.} have designed a very simple range analysis,
which does not handle comparisons between variables, for instance.
They also do not discuss the precision of their implementation, but only its
runtime.

% Our contribution.
In this paper we provide a complete description of a range analysis algorithm,
and show extensive experimental data that justifies our engineering choices.
Our first algorithmic contribution on top of previous works is a three-phases
approach to handle comparisons between variables without resorting to any
exponential time technique.
The few publicly available implementations of range analyses that we are
aware of, such as those in FLEX~\footnote{The MIT's FLEX/Harpoon compiler
provides an implementation of Stephenson's algorithm~\cite{Stephenson00}, and
is available at \texttt{http://flex.cscott.net/Harpoon/}.} and
gcc~\footnote{Gcc's VRP pass implements a variant of Patterson's
algorithm~\cite{Patterson95}.} only deal with comparisons between variables and
constants.
Even theoretical works, such as Su and Wagner's~\cite{Su05} or Gawlitza
{\em et al.}'s~\cite{Gawlitza09} suffer from this limitation.
This deficiency is one of the reasons explaining why none of these works has
made their way into industrial-strength compilers.
Two other insights allow our implementation to scale up to very large
programs.
We use Bodik's Extended Static Single Assignment (e-SSA) form~\cite{Bodik00} to
perform range analysis sparsely.
This program representation ensures that the interval associated with
a variable is constant along its entire live range, and allows us to compute
path-sensitive results.
Finally, we process the strongly connected components that underline our
constraint system in topological ordering.
It is well-known that this technique is essential to speedup constraint solving
algorithms~\cite[Sec 6.3]{Nielson99}; however, we show that a careful
propagation of information along strong components not only gives us speed, but
also improves the precision of our results.

% The history of our contribution.
This work concludes a two years long effort to produce a solid and scalable
implementation of range analysis.
Our first endeavor to implement such an algorithm was based on Su and
Wagner's constraint system, which can be solved exactly in polynomial
time~\cite{Su04,Su05}.
However, although we could use their formulation to handle a subset of C-like
constructs, their description of how to deal with loops was
not very explicit.
Thus, in order to solve loops we adopted Gawlitza
{\em et al.}'s~\cite{Gawlitza09} approach.
This technique uses the Bellman-Ford algorithm to detect increasing or
decreasing cycles in the constraint system, and then saturates these cycles
via a simple widening operator.
A detailed description of our implementation has been published by
Couto and Pereira~\cite{Couto11}.
Nevertheless, the inability to handle comparisons between variables, and the
cubic complexity of the Bellman-Ford method eventually lead us to seek
alternative solutions to range analysis.
This quest reached a pinnacle in the present work.

% Our experimental numbers.
We have implemented our algorithm in the LLVM compiler~\cite{Lattner04}, and
have used it to analyze a test suite with 2.72 million lines of C code.
In Section~\ref{sec:exp} we provide empirical evidence showing that our
implementation is fast: it globally analyzes the gcc source code in less than
15 seconds, for instance.
It is also precise: our results are similar to Stephenson
{\em et al}'s~\cite{Stephenson00}, even though out analysis does not require
a backward propagation phase.
Furthermore, we have been able to find tight bounds to the majority of the
examples used by Costan {\em et al.}~\cite{Costan05} and Lakhdar
{\em et al.}~\cite{Lakhdar11}, who rely on much more costly methods.

% The rest of this paper.

\section{Background}
\label{sec:bck}

% Define the lattice, the constraints and the valuation I.
Following Gawlitza {\em et al.}'s notation, we shall be performing arithmetic
operations over the complete lattice
$\cal{Z} = \mathbb{Z} \cup \{-\infty, +\infty\}$, where the ordering is
naturally given by $-\infty < \ldots < -2 < -1 < 0 < 1 < 2 < \ldots +\infty$.
For any $x > -\infty$ we define:

\begin{tabular}{lcl}
$x + \infty = \infty, x \neq -\infty$ & \mbox{\hspace{0.1cm}} & $x - \infty = - \infty, x \neq +\infty$ \\
$x \times \infty = \infty$ if $x > 0$ & & $x \times \infty = -\infty$ if $x < 0$ \\
$0 \times \infty = 0$ & & $(-\infty) \times \infty = \ $ not defined $$ \\
\end{tabular}

From the lattice $\varset{Z}$ we define the product lattice
$\varset{Z}^2$, which is partially ordered by the subset relation
$\sqsubseteq$.
$\varset{Z}^2$ is defined as follows:

\begin{equation*}
\varset{Z}^2 = \emptyset \cup \{[z_1, z_2] | \ z_1,z_2 \in \varset{Z},
\ z_1 \leq z_2, \  -\infty < z_2 \}
\end{equation*}

Given an interval $\iota = [l, u]$, we let $\lb{\iota} = l$, and
$\ub{\iota} = u$.
We let \varset{V} be a set of constraint variables, and
$I: \varset{V} \mapsto \varset{Z}^2$ a
mapping from these variables to intervals in $\varset{Z}^2$.
Our objective is to solve a constraint system $C$, formed by constraints such
as those seen in Figure~\ref{fig:eval_function}(left).
Figure~\ref{fig:eval_function}(right) defines a valuation function $e$ that
computes $Y = f(\ldots)$, given I.
Armed with these concepts, we define the range analysis problem as follows:

\begin{definition}
\label{def:rcp}
\textsc{Range Analysis Problem} \\
\textbf{Input:} a set $C$ of constraints ranging over a set \varset{V} of
variables. \\
\textbf{Output:} a mapping I such that, for any variable
$V \in \varset{V}$, e(V) = I[V].
\end{definition}

\begin{figure}[hbt]
\begin{small}
\begin{eqnarray*}
\begin{array}{r@{\hspace{0.5cm}}c}
Y = [l, u]
&
e(Y) = [l, u]
\\
\\
Y = \phi (X_1, X_2)
&
\inferrule{I[X_1]=[l_1, u_1] \\ I[X_2]=[l_2, u_2]}
{e(Y) = [\mbox{min}(l_1, l_2), \mbox{max}(u_1, u_2)]}
\\
\\
Y = X_1 + X_2
&
\inferrule{I[X_1]=[l_1, u_1] \\ I[X_2]=[l_2, u_2]}
{e(Y) = [l_1 + l_2, u_1 + u_2]}
\\
\\
Y = X_1 \times X_2
&
\inferrule{I[X_1]=[l_1, u_1] \\ I[X_2]=[l_2, u_2] \\ L = \{l_1l_2, l_1u_2, u_1l_2, u_1u_2\}}
{e(Y) = [\mbox{min}(L), \mbox{max}(L)]}
\\
\\
Y = aX + b
&
\inferrule{I[X]=[l, u] \\ k_l = al + b \\ k_u = au + b}
{e(Y) = [\mbox{min}(k_l, k_u), \mbox{max}(k_l, k_u)]}
\\
\\
Y = X \sqcap [l', u']
&
\inferrule{I[X]=[l, u]}
{e(Y) \leftarrow [\mbox{max}(l, l'), \mbox{min}(u, u')]}
\end{array}
\end{eqnarray*}
\caption{\label{fig:eval_function}
A suite of constraints that produce an instance of the range analysis problem.}
\end{small}
\end{figure}

We will use the program in Figure~\ref{fig:ex1}(a) as the running example
to illustrate our range analysis.
Figure~\ref{fig:ex1}(b) shows the same program in SSA form~\cite{Cytron91},
and Figure~\ref{fig:ex1}(c) outlines the constraints that we extract from this
program.
There is a clear correspondence between instructions and constraints.
A possible solution to the range analysis problem, as obtained via the
techniques that we will introduce in Section~\ref{sec:algo}, is given in
Figure~\ref{fig:ex1}(d).
The SSA form, so conspicuous in modern compilers, leads to a very conservative
solution.
As we will see shortly, we can improve this solution substantially by using
a more sophisticated program representation -- the e-SSA form -- which
gives us path-sensitiveness.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=1\textwidth]{images/ex1}
\end{center}
\caption{\label{fig:ex1}
(a) Example program.
(b) Control Flow Graph in SSA form.
(c) Constraints that we extract from the program.
(d) Possible solution to the range analysis problem.}
\end{figure}

\section{Our Design of a Range Analysis Algorithm}
\label{sec:algo}

% Program representation
% Extracting constraints
% Building the Constraint graph (pseudo-edges)
% Macro algorithm
% Micro algorithm

In this section we explain the algorithm that we use to solve the range
analysis problem.
This algorithm involves a number of steps.
First, we convert the program to a suitable intermediate representation that
makes it easier to extract constraints.
From these constraints, we build a dependence graph that allows us to do
range analysis sparsely.
Finally, we solve the constraints applying different fix-point iterators on
this dependence graph.
Figure~\ref{fig:algorithm} gives a global view of this algorithm.
Some of the steps in the algorithm are optional.
They improve the precision of the range analysis, at the expense of a longer
running time.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=1\textwidth]{images/algorithm}
\end{center}
\caption{\label{fig:algorithm}
Our implementation of range analysis.
Rounded boxes are optional steps.}
\end{figure}

\noindent
\textbf{Choosing a Program Representation.}
The solution to the range analysis problem in Figure~\ref{fig:ex1}
is imprecise because we did not take conditional tests into considerations.
Branches give us information about the ranges that some variables assume, but
only at {\em specific} program points.
For instance, given a test such as $(k_1 < 100)?$ in  Figure~\ref{fig:ex1}(b),
we know that $I[k_1] \sqsubseteq [-\infty, 99]$ whenever the condition is true.
In order to encode this information, we might split the live range of $k_1$
right after the branching point; thus, creating two new variables, one at the
path where the condition is true, and another where it is false.
There is a program representation, introduced by Bodik
{\em et al.}~\cite{Bodik00}, that performs this live range splitting:
the {\em Extended Static Single Assignment} form, or e-SSA for short.

Given that the exact rules to convert a program to e-SSA form have never been
explicitly stated in the literature, we describe our rules as follows.
Let $(v < c)?$ be a conditional test, and let $l_t$ and $l_f$ be labels in
the program, such that $l_t$ is the target of the test if the condition is true,
and $l_f$ is the target when the condition is false.
We split the live range of $v$ at any of these points if at least one of two
conditions is true:
(i) $l_f$ or $l_t$ dominate any use of $v$;
(ii) there exist a use of $v$ at the dominance frontier of $l_f$ or $l_t$.
For the notions of dominance and dominance-frontier, see Aho
{\em et al.}~\cite[p.656]{Aho06}.
To split the live range of $v$ at $l_f$ we insert at this
program point a copy $v_f = v \sqcap [c, +\infty]$, where $v_f$ is a fresh name.
We then rename every use of $v$ that is dominated by $l_f$ to $v_f$.
Dually, if we must split at $l_t$, then we create at this point a copy
$v_t = v \sqcap [-\infty, c-1]$, and rename variables accordingly.
If the conditional uses two variables, e.g., $(v_1 < v_2)?$, then we create
intersections bound to {\em futures}.
We insert, at $l_f$, $v_{1f} = v_1 \sqcap [\fun{ft}(v_2), +\infty]$,
and $v_{2f} = v_2 \sqcap [-\infty, \fun{ft}(v_1)]$.
Similarly, at $l_t$ we insert
$v_{1v} = v_1 \sqcap [-\infty, \fun{ft}(v_2) - 1]$
and $v_{2f} = v_2 \sqcap [\fun{ft}(v_1) + 1, +\infty]$.

We use the notation $\fun{ft}(v)$ to denote the {\em future} bounds of a
variable.
As we will show in Section~\ref{sub:micro}, once the growth pattern of $v$ is
known, we can replace $\fun{ft}(v)$ by an actual value.
Once we are done placing copies, we insert $\phi$-functions into the
transformed program to convert it to SSA form.
This last step avoids that two different names given to the same original
variable be simultaneously alive at the program code.
Figure~\ref{fig:ex_standard_eSSA}(a) shows our running example changed into
standard e-SSA form.
The part (b) of this figure shows the solution that we get to this new
program.
The e-SSA form allows us to bind interval information directly to the live
ranges of variables; thus, giving us the opportunity to solve range analysis
sparsely.
More traditional approaches, which we call {\em dense analyses}, bind
interval information to pairs formed by variables and program points.

% TODO: check with Victor ** eSSA description could also say that we create sigmas even if lt/lf does not dominate any use of v (under some conditions), i.e. when there is a use in dominance frontier inside a phi function **

\begin{figure}[t!]
\begin{center}
\includegraphics[width=0.9\textwidth]{images/ex_standard_eSSA}
\end{center}
\caption{\label{fig:ex_standard_eSSA}
(a) The control flow graph from Figure~\ref{fig:ex1}(b) converted to standard
e-SSA form.
(b) A solution to the range analysis problem
}
\end{figure}

%The standard e-SSA form serves well analysis that need range information
%at the use site of variables, such as conditional constant propagation,
%redundant code elimination and detection of buffer overflows.
%On the other hand, there exist compilation algorithms that require range
%information at the whole live range of variables.
%Examples of such algorithms include the family of
%bitwidth aware register allocators~\cite{Barik06,Pereira08,Tallam03} and
%hardware synthesizers~\cite{Cong05,Mahlke01,Stephenson00}.
%In order to provide extra precision to these algorithms, we may work with a
%non-pruned version of e-SSA form, which we produce by simply inserting copies
%after every conditional test, and then reconverting the entire program to
%SSA form.
%Figure~\ref{fig:ex_non_pruned} illustrates the difference between these two
%program representations.
%Notice that the non-pruned flavor might introduce dead variables in the
%program code, which can be removed by standard dead code elimination.

%\begin{figure}[t!]
%\begin{center}
%\includegraphics[width=0.9\textwidth]{images/ex_non_pruned}
%\end{center}
%\caption{\label{fig:ex_non_pruned}
%(a) Example program.
%(b) Standard e-SSA form.
%(c) Non-pruned e-SSA form.}
%\end{figure}

\noindent
\textbf{Extracting Constraints.}
Our implementation handles 18 different assembly instructions.
The constraints in Figure~\ref{fig:eval_function} show only a few examples.
Instructions that we did not show include, for instance, the multiplicative
operators \texttt{div} and \texttt{modulus},
the bitwise operators \texttt{and}, \texttt{or}, \texttt{xor} and \texttt{neg},
the different types of shifts, and the
logical operators \texttt{andalso}, \texttt{orelse} and \texttt{not}.
Most of these instructions are sign-agnostic; that is, given that numbers are
internally represented in 2's complement, the same implementation of a
constraint handles positive and negative numbers.
However, there are instructions that require different constraints, depending on
the input being signed or not.
Examples include \texttt{modulus} and \texttt{div}.
We also handle different kinds of type conversion, e.g., converting 8-bit
characters to 32-bit integers and vice-versa.
In addition to constraints that represent actual assembly instructions, we have
constraints to represent $\phi$-functions, and intersections, as seen in
Figure~\ref{fig:eval_function}.
The growth analysis that we will introduce in Section~\ref{sub:micro}
require monotonic transfer functions.
Many assembly operations, such as modulus or division, do not afford us
this monotonicity.
However, these non-monotonic instructions have conservative
approximations~\cite{Warren02}.

%Figure~\ref{fig:constraints} shows some examples of constraints that we
%derive from typical assembly instructions.
%
%\begin{figure}[t!]\begin{center}
%\begin{tabular}{|l@{\hspace{0.2cm}}l@{\hspace{0.3cm}}l|} \hline
%\textbf{Description} & \textbf{Operation} & \textbf{Constraint} \\ \hline
%Constant & $v = c$ & $V = [c, c]$ \\
%Assignment & $v_1 = v_2$ & $V_1 = V_2$ \\
%Multiplication & $v_1 = v_2 * c$ & $V_1 = c V_2$ \\
%Int. division & $v_1 = v_2 \ / \ v_3$ & $V_1 = V_2; V_1 = - V_2$ \\
%Modulus & $v_1 = v_2 \ \% \ c$ & $V_1 = [0, c - 1]$ \\
%Bitwise and & $v_1 = v_2 \ \& \ c$ & $V_1 = v_2 \sqcap [0, c]$ \\
%Bitwise or & $v_1 = v_2 \ | \ v_3$ & $V_1 = V_2 + V_3$ \\
%Left shift & $v_1 = v_2 \ \qside \ c$ & $V_1 = 2^c V_2$ \\ \hline
%\end{tabular}
%\end{center}
%\caption{\label{fig:constraints}Example of constraint derivation rules.
%We let $c$ denote a positive constant, $\{v, v_1, v_2\}$ are program
%variables and $\{V, V_1, V_2\} \subset \cal{V}$.}
%\end{figure}

\noindent
\textbf{The Constraint Graph.}
The main data structure that we use to solve the range analysis problem is
a variation of Ferrante {\em et al.}'s {\em program dependence
graph}~\cite{Ferrante87}.
For each constraint variable $V$ we create a variable node $N_v$.
For each constraint $C$ we create a constraint node $N_c$.
We add an edge from $N_v$ to $N_c$ if the name $V$ is used in $C$.
We add an edge from $N_c$ to $N_v$ if the constraint $C$ defines the name
$V$.
Figure~\ref{fig:ex_graph} shows the dependence graph that we build for the
e-SSA form program given in Figure~\ref{fig:ex_standard_eSSA}(a).
If $V$ is used by $C$ as the input of a future, then the edge from
$N_v$ to $N_c$ represents what Ferrante {\em et al.} call a {\em control
dependence}~\cite[p.323]{Ferrante87}.
We use dashed lines to represent these edges.
All the other edges denote {\em data dependences}~\cite[p.322]{Ferrante87}.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=0.7\textwidth]{images/ex_graph}
\end{center}
\caption{\label{fig:ex_graph}
The dependence graph that we build to the program in
Figure~\ref{fig:ex_standard_eSSA}.}
\end{figure}

\noindent
\textbf{Strongly Connected Components.}
To solve range analysis we find all the strongly connected components (SCCs) of
the dependence graph and collapse them into single nodes; thus, obtaining a
directed acyclic graph.
We then sort the resulting DAG topologically, and apply the analyses from
Section~\ref{sub:micro} on every SCC in topological order.
Once we solve the range analysis problem for a SCC, we propagate the
intervals that we found to the variable nodes at the {\em frontier} of this
SCC.
A variable node $N_v$ is said to be in the frontier of a strongly connected
component $S$ if:
(i) $N_v \notin S$; and
(ii) there exists a variable node $N_v' \in S$, and a constraint node $N_c$,
such that $N_v \leftarrow N_c$, and $N_c \leftarrow N_v'$.
This propagation ensures that when analyzing a strongly connected component $S$
any influence that $S$ might suffer from nodes outside it has already been
considered.

When finding strongly connected components, we take control dependence
edges into consideration.
For instance, in Figure~\ref{fig:ex_graph} the nodes that correspond to the
variables $i_1$, $i_2$, $i_t$, $j_1$, $j_2$ and $j_t$ form a single component.
The dashed edges, which represent control dependences, keep all these variables
connected.
In this way, we ensure that, upon stumbling upon an interval associated with
future bounds, e.g., $\fun{ft}(v)$, either variable $v$ has been solved in a
previous component, or it belongs to the current component.
In the latter case, as we will see soon, we can still take $v$'s interval into
consideration.
As we show in Section~\ref{sec:exp}, most of the strong components in actual
programs are singletons.
Furthermore, the composite components tend to be small.
These two facts ensure that the more costly parts of our algorithm only have to
handle small inputs.

\subsection{Finding Ranges in Strongly Connected Components}
\label{sub:micro}

Given a strongly connected component of the dependence graph with $N$ nodes,
we solve the range analysis problem in three-steps:
\begin{enumerate}
\item Run growth analysis: $O(N)$.

\item Fix intersections: $O(N)$.

\item Apply the narrowing operator: $O(N^2)$.

\end{enumerate}
However, before we start, we remove control dependence edges from the
strongly connected component, as they have no semantics to our transfer
functions.

\noindent
\textbf{Growth Analysis.}
The first step of our algorithm consists in determining how the interval bound
to each variable grows.
We ensure termination of this phase via Cousot and Cousot's widening
operator~\cite[p.247]{Cousot77}.
The possible behaviors of an interval are:
(i) does not change;
(ii) grows towards $+\infty$;
(iii) grows towards $-\infty$; and
(iv) grows in both directions.
The lattice of abstract states, plus a constraint system representing our
growth analysis is given in Figure~\ref{fig:growth_analysis}.
Because the lattice has height three, the intervals bound to each variable can
change at most three times.
However, as we show in Figure~\ref{fig:essaDivSSA}(Bottom), we can be more
precise if we evaluate the constraints a few times before applying widening.

\begin{figure}[t!]
\begin{center}
\begin{tabular}{c@{\hspace{1.5cm}}c}
\begin{minipage}{2cm}
\includegraphics{images/growth_lattice}
\end{minipage}
&
\begin{minipage}{8cm}
\begin{small}
\begin{eqnarray*}
\begin{array}{c@{\hspace{0.5cm}}c}
\inferrule{I[Y] = [\bot, \bot]}{I[Y] \leftarrow e(Y)}
&
\inferrule{\lb{e(Y)} < \lb{I[Y]} \\ \ub{e(Y)} > \ub{I[Y]}}
{I[Y] \leftarrow [-\infty, +\infty ]}
\\
\\
\inferrule{\lb{e(Y)} < \lb{I[Y]}}
{I[Y] \leftarrow [-\infty, \ub{I[Y]}]}
&
\inferrule{\ub{e(Y)} > \ub{I[Y]}}
{I[Y] \leftarrow [\lb{I[Y]}, +\infty]}
\end{array}
\end{eqnarray*}
\end{small}
\end{minipage}
\end{tabular}
\end{center}
\caption{\label{fig:growth_analysis}
(Left) The lattice of the growth analysis.
(Right) Cousot and Cousot's widening operator. We evaluate the rules from
left-to-right, top-to-bottom, and stop upon finding a pattern matching.
Again: given an interval $\iota = [l, u]$, we let $\lb{\iota} = l$, and
$\ub{\iota} = u$}
\end{figure}

\noindent
\textbf{Fixing futures.}
The ranges found by the growth analysis tells us which variables have fixed
bounds, independent on the intersections in the constraint system.
Thus, we can use actual limits to replace intersections bounded by futures.
Figure~\ref{fig:fix_intersects} shows the rules to perform these substitutions.
In order to correctly replace a future $\fun{ft}(V)$ that limits a variable
$V'$, we need to have already applied the growth analysis onto $V$.
Had we considered only data dependence edges, then it would be possible
that $V'$ be analyzed before $V$.
However, because of control dependence edges, this case cannot happen.
The control dependence edges ensure that any topological ordering of the
constraint graph either places $N_v$ before $N_{v'}$, or places these nodes
in the same strongly connected component.
For instance, in Figure~\ref{fig:ex_graph}, variables $J_1$ and $I_t$ are in
the same SCC only because of the control dependence edges.

\begin{figure}[t!]
\begin{center}
\begin{eqnarray*}
\begin{array}{c}
\inferrule{Y = X \sqcap [l, \fun{ft}(V) + c] \\ \ub{I[V]} = u}
{Y = X \sqcap [l, u + c]} \mbox{\hspace{0.3cm}} u, c \in \mathbb{Z} \cup \{-\infty, +\infty\}
\\
\\
\inferrule{Y = X \sqcap [\fun{ft}(V) + c, u] \\ \lb{I[V]} = l}
{Y = X \sqcap [l + c, u]} \mbox{\hspace{0.3cm}} l, c \in \mathbb{Z} \cup \{-\infty, +\infty\}
\end{array}
\end{eqnarray*}
\end{center}
\caption{\label{fig:fix_intersects}Rules to replace futures by actual
bounds. $S$ is the interval bound to each variable after the widening
analysis.}
\end{figure}

%\begin{figure}[t!]
%\begin{equation*}
%S[X] \leftarrow
%\begin{cases}
%``\ ? \ " \ \ \mbox{if} \ I[X] = [-\infty, +\infty]\\
%``\downarrow" \ \mbox{if} \ I[X] = [-\infty, c], c \in \mathbb{Z} \\
%``\uparrow" \ \mbox{if} \ I[X] = [c, +\infty],  c \in \mathbb{Z} \\
%`` \ 0 \ " \ \ \mbox{if} \ I[X] = [c_1, c_2], c_1, c_2 \in \mathbb{Z}
%\end{cases}
%\end{equation*}
%\caption{\label{fig:st_mem}The growth state of each variable.}
%\end{figure}

\noindent
\textbf{Narrowing Analysis.}
The growth analysis associates very conservative bounds to each variable.
Thus, the last step of our algorithm consists in narrowing these intervals.
We accomplish this step via Cousot and Cousot's classic narrowing
operator~\cite[248]{Cousot77}, which we show in
Figure~\ref{fig:crop_analysis}.

\begin{figure}[t!]
\begin{center}
\begin{eqnarray*}
\begin{array}{c@{\hspace{0.9cm}}c}
\inferrule{\lb{I[Y]} = -\infty \\ \lb{e(Y)} > -\infty}
{I[Y] \leftarrow [\lb{e(Y)}, \ub{I[Y]}]}
&
\inferrule{\lb{I[Y]} > \lb{e(Y)}}
{I[Y] \leftarrow [\lb{e(Y)}, \ub{I[Y]}]}
\\
\\
\inferrule{\ub{I[Y]} = +\infty \\ \ub{e(Y)} < +\infty}
{I[Y] \leftarrow [\lb{I[Y]}, \ub{e(Y)}]}
&
\inferrule{\ub{I[Y]} < \ub{e(Y)}}
{I[Y] \leftarrow [\lb{I[Y]}, \ub{e(Y)}]}
\end{array}
\end{eqnarray*}
\end{center}
\caption{\label{fig:crop_analysis}Cousot and Cousot's narrowing operator.}
\end{figure}

\noindent
\textbf{Example:}
Continuing with our example, Figure~\ref{fig:ex_partition_grow_crop} shows
the application of our algorithm on the last strong component of
Figure~\ref{fig:ex_graph}.
Upon meeting this SCC, we have already determined that the interval
$[0, 0]$ is bound to $I_0$ and that the interval $[100, 100]$ is bound to
$J_0$, as we show in Figure~\ref{fig:ex_graph}(a).
We are not guaranteed to find the least fix point of a constraint system.
However, in this example we did it.
We emphasize that finding this tight solution was only possible because of
the topological ordering of the constraint graph in
Figure~\ref{fig:ex_graph}.
Had we applied the widening operator onto the whole graph, then we would
have found out that variable $j_0$ is bound to $[-\infty, +\infty]$,
because
(i) it receives its interval directly from variable $k_t$, which is upper
bounded by $+\infty$, and
(ii) it is part of a negative cycle.
On the other hand, because we only analyze $j$'s SCC after we have
analyzed $k$'s, $k$ only contribute the constant range $[0, 99]$ to $j_0$.

Figure~\ref{fig:ex_standard_eSSA}(b) shows our final solution for this example.
This solution is very precise, in the sense that it is the maximum fixed
point of the constraint system given in Figure~\ref{fig:eval_function}.
However, the solution is still an over approximation of the dynamic behavior of
the program in Figure~\ref{fig:ex_standard_eSSA}.
For instance, we have found that variable $i$ could reach the upper value of
99.
In any actual run of the program, $i$ could be at most 50.
Analysis on relational lattices, such as the polyhedron~\cite{Cousot78} or the
Octagon~\cite{Mine06} domains, can infer such tighter bounds, as shown by
Lakhdar-Chaouch {\em et al.}~\cite{Lakhdar11}.
However, analyses on these higher dimensional domains are much more
computationally expensive than analyses on the interval domain~\cite{Oh12}.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=\textwidth]{images/ex_partition_grow_crop}
\end{center}
\caption{\label{fig:ex_partition_grow_crop}
Four snapshots of the last SCC of Figure~\ref{fig:ex_standard_eSSA}.
(a) After removing control dependence edges.
(b) After running the growth analysis.
(c) After fixing the intersections bound to futures.
(d) After running the narrowing analysis.}
\end{figure}

%\subsection{Alternative Narrowing Operators}
%\label{sub:alt}
%
%Cousot and Cousot's narrowing operator, given in
%Figure~\ref{fig:crop_analysis} is not guaranteed to find the least fixpoint
%of a system of constraints.
%In order to improve the bounds, trading time for precision, we have
%experimented with two other operators.
%The first, is the evaluation function itself, which is given in
%Figure~\ref{fig:eval_function}.
%By iterating it until reaching a fixpoint, we are guaranteed to find the
%optimal solution to the narrowing analysis.
%However, this approach is not practical, because it is equivalent to
%concretely interpreting the program.
%The second alternative that we have tested consists in applying the
%evaluation function on each constraint at most once per intersection.
%We call this approach {\em depth-first narrowing}, because, starting from
%an intersection we visit every node that might be influenced by it.
%After removing control dependence edges, if we are left with a strongly
%connected component containing only {\em ascencing constraints}, that
%is, constraints that only increase the upper bounds of intervals, then this
%approach is guaranteed to find the least fixpoint.
%The same is true if we have a SCC with only {\em descending constraints}.

\section{Experiments}
\label{sec:exp}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We have implemented our range analysis algorithm in LLVM 3.0, and have run
experiments on a Intel quad core CPU with a 2.40GHz clock, and 3.6GB of RAM.
Each core has a 4,096KB L1 cache.
We use Linux Ubuntu 10.04.4.
Our implementation of range analysis has 3,958 lines of
commented C++ code, and our e-SSA conversion module has 672 lines.
We have analyzed 428 C programs that constitute the LLVM test suite plus the
integer benchmarks in SPEC CPU 2006.
Together, these programs contain 4.76 million assembly instructions.

\noindent
\textbf{Structural Properties.}
Table~\ref{tab:struct} shows data taken from the SPEC 2006 integer benchmark
suite for the inter-procedural version of our analysis.
Three observations provide evidence that our algorithm is linear in practice,
as we will show in Figure~\ref{fig:TimeCorr}.
First, most of the strongly connected components found in constraint graphs are
singletons.
For instance, 99.11\% of the SCCs in gcc (\texttt{403.gcc}) have only
one node.
Hence, the most complex phases of our algorithm are confined to a small part
of the constraint graph.
Second, the strongly connected components are not large -- our largest
component, found in \texttt{403.gcc}, has 2,131 nodes.
It exists due to a long chain of mutually recursive function calls.
We do not count control dependency edges when measuring the size of SCCs, as
they have no impact on the algorithm explained in Section~\ref{sub:micro}.
Third, as we see in the table, any node was visited at most three times by
the narrowing operator.

\begin{table}[t!]
\begin{center}
\begin{scriptsize}
\renewcommand{\tabcolsep}{0.1cm}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|} \hline
Name	    & 464 &	473 & 483 & 458 & 429 & 471 & 403 & 445 & 462 & 401 & 456 \\ \hline
\#I  & 139626 & 8900 & 593765 & 26947 & 2738 & 94036 & 887191 & 141330 &  6237 & 17663 & 61736 \\ \hline
\#$N_v$ & 97494 & 5490 & 352423 & 14407 & 1853	 & 50095 & 449442 & 84846 & 3293 & 12517 & 38409 \\ \hline
\#$N_c$ & 69530 & 3799 & 213535 & 11164 & 1156 & 28943 & 309310 & 63506 & 2336 & 8731 & 24530 \\ \hline
\#S  & 91731 & 5108 & 346451 & 13309 & 1751 & 48784 & 435372 & 79909 & 3003 & 11539 & 35619 \\ \hline
L    & 38 & 49 & 120 & 116 & 16 & 258 & 2131 & 311 & 9 & 18 & 41  \\ \hline
\#U  & 89461 & 4983 & 344132 & 12919 & 1726 & 48400 & 431703 & 78407 & 2885 & 11267 & 34631 \\ \hline
\#V  & 3	& 2 & 2 & 2 & 2 & 2 & 2 & 3 & 2 & 2 & 2 \\ \hline
\end{tabular}
\end{scriptsize}
\end{center}
\caption{\label{tab:struct}Structural Properties of the programs in the SPEC CPU
2006 integer benchmark suite.
\#I: number of assembly instructions in the original program.
\#$N_v$: number of variable nodes in the constraint graph.
\#$N_c$: number of constraint nodes in the constraint graph.
% TODO: add also number of constraint nodes.
\#S: number of strongly connected components (SCC).
L: size of the largest SCC.
\#U: number of SCCs with only one node.
\#V: maximum number of times any variable node was visited by narrowing.}
\end{table}%

% Absolute Time and complexity
% =============================

\noindent
\textbf{Asymptotic complexity.}
Figure~\ref{fig:TimeCorr} provides a visual comparison between the time to
run our algorithm and the size of the input programs.
We show data for the 100 largest benchmarks in our test suite, in number
of variable nodes in the constraint graph.
We perform function inlining before running our analysis, to increase program
sizes.
Each point in the X line corresponds to a benchmark.
We analyze the smallest benchmark in this set, \texttt{Prolangs-C}, which has
1,131 variable nodes in the constraint graph, in 20ms.
We take 15,91secs to analyze our largest benchmark, \texttt{403.gcc}, which,
after function inlining, has 1,266,273 assembly instructions, and a
constraint graph with 679,652 variable nodes.
For this data set, the coefficient of determination $(R^2)$ is 0.967, which
provides very strong evidence about the linear asymptotic complexity of our
implementation.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=0.9\textwidth]{images/TimeCorr}
%\includegraphics[width=0.4\textwidth]{images/Complexity}
\end{center}
\caption{\label{fig:TimeCorr}
Correlation between program size (measured in number of nodes in constraint
graphs after inlining) and analysis runtime (ms).
Coefficient of determination = 0.967.
}
\end{figure}

\noindent
\textbf{Time distribution.}
Figure~\ref{fig:timeComp}(Left) shows the distribution of processing time
among the main phases of the inter-procedural version of our algorithm.
Functions are inlined to increase program sizes.
We analyze the SPEC 2006 integer benchmarks in 41.42secs.
The time to build the e-SSA representation is 7.03secs; the time to extract
constraints from the program and build the constraint graphs is 7.87secs;
the time to find SCCs via Nuutila's algorithm and sort them topologically is
10.55secs, and the time to run the analysis itself is 15.96secs.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=0.48\textwidth]{images/timeDistr}
\includegraphics[width=0.48\textwidth]{images/timeIntraInterInline}
\end{center}
\caption{\label{fig:timeComp}
(Left) Time distribution across different phases of our inter+inline
algorithm.
(Right) Runtime comparison between the intra, inter e inter+inline versions of
our algorithm.
The bars are normalized to the time to run the intra-procedural version.
}
\end{figure}

\noindent
\textbf{Intra vs Inter-procedural runtimes.}
As we saw in Figure~\ref{fig:algorithm}, our implementation supports
intra- and inter-procedural execution modes.
On top of these modes, we can optionally perform function inline to obtain
limited context-sensitiveness.
Our baseline compiler, LLVM, is able to inline non-recursive function calls.
Figure~\ref{fig:timeComp}(Right) compares three different execution modes.
Times are normalized to the time to run the intra-procedural analysis
without inlining.
On the average, the intra-procedural mode is 28.92\% faster than the
inter-procedural mode.
If we perform function inlining, then this difference is larger: 45.87\%.
These numbers are close because we run our analysis on strong components.

\noindent
\textbf{The impact of strong components in time and precision.}
Figure~\ref{fig:impactEssaSCC}(a) shows that strong components dramatically
decrease the runtime of our analysis.
For instance, this design improves the analysis time of \texttt{483.xalancbmk}
452 times!
Moreover, as the example in Section~\ref{sub:micro} shows, our algorithm benefits
from the partitioning of constraint graphs into SCCs to be more precise.
Figure~\ref{fig:impactEssaSCC}(b) measures this precision.
Range propagation along strong components allows us, for example, to reduce
41.5\% more bits in \texttt{401.bzip}.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=\textwidth]{images/impactEssaSCC}
\end{center}
\caption{\label{fig:impactEssaSCC}
How strong components and the e-SSA representation improve our algorithm.
This data refer to the inter-procedural analysis with function inlining.
}
\end{figure}

\noindent
\textbf{Impact of the e-SSA on runtime.}
Figure~\ref{fig:impactEssaSCC}(c) shows that the e-SSA conversion increases
the size of the SPEC 2006 integer benchmarks by 6.52\%.
This growth is due to the $\sigma$-functions used to split live ranges
after conditionals, and to the extra $\phi$-functions necessary to preserve the
single static assignment property.
As a consequence of the program growth, the e-SSA form increases the runtime of
our analysis by 7.11\%, as seen in Figure~\ref{fig:impactEssaSCC}(d).
We are comparing only the time to perform range analysis.
The bars do not include the time to perform the e-SSA conversion, which we
showed in Figure~\ref{fig:timeComp}(Left).

\noindent
\textbf{The impact of whole program analysis on precision.}
Figure~\ref{fig:essaDivSSA}(Top) shows that the whole program analysis increases
moderately the precision of our analysis.
We show results for the five largest programs in three different categories of
benchmarks: SPEC CPU 2006, the Stanford Suite~\footnote{\texttt{http://classes.engineering.wustl.edu/cse465/docs/BCCExamples/stanford.c}} and
Bitwise~\cite{Stephenson00}.
Our initial goal when developing this analysis was to support a bitwidth-aware
register allocator.
Thus, we measure precision by the average number of bits that our
analysis allows us to save per program variables.
It is very important to notice that we do not consider the constants in the
program text in our statics of precision.
In other words, we only measure bitwidth reduction in variables that a constant
propagation step could not remove.
Our results for the SPEC programs were disappointingly: on the average, the
intra-procedural version of our analysis saves 5.23\% of the total program
bitwidth.
By using the inter-procedural version, with function inlining to simulate
context-sensitiveness, we can increase this number to 8.89\%.
A manual inspection of the SPEC programs reveal that this result is expected:
they manipulate files, and their code do no provide enough
explicit constants to power our analysis up.
However, with numerical benchmarks we fare much better.
On the average our inter-procedural algorithm reduces the bitwidth of the
Stanford benchmarks by 36.24\%.
Notice that this number is better than the one we get with inlining (31.97\%),
because functions in those benchmarks are called only at one location.
Finally, for Bitwise we obtain a bitwidth reduction of 12.27\%.
However, this average is brought down by two outliers: \texttt{edge\_detect} and
\texttt{sha}, which cannot be reduced.
The bitwise benchmarks were implemented by Stephenson
{\em et al.}~\cite{Stephenson00} to validate their bitwidth analysis.
Our results are on par with those found by the original authors.
The bitwise programs contain only the \texttt{main} function; thus, different
versions of our algorithm find the same results.

\noindent
\textbf{The impact of e-SSA on precision.}
Figure~\ref{fig:essaDivSSA}(Middle) shows that the e-SSA transformation improves
the precision of our analysis dramatically.
If we convert the input programs from the original SSA representation used
by LLVM to the e-SSA format, then, in some cases, e.g.,
\texttt{Stanford.queens}, we can increase bitwidth reduction by 154x.
This impressive difference happens because the standard SSA representation does
not give our analysis any subsidy to acquire information from the outcome of
conditionals.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=1\textwidth]{images/essaDivSSA}
\end{center}
\caption{\label{fig:essaDivSSA}
The impact of different design trade-offs on precision of range analysis.}
\end{figure}

\noindent
\textbf{The impact of the number of iterations before widening on precision.}
It is well-known that premature widening might cause loss of
information~\cite[Sec5.2]{Cousot09}.
Figure~\ref{fig:essaDivSSA} shows how the precision of our analysis varies with
the number of iterations used before resorting to widening.
We have applied the widening operator in Figure~\ref{fig:growth_analysis}
after each node was visited 1, 2 and 16 times.

\section{Related works}
\label{sec:rel}

% Early implementations of range analysis.
Range analysis is an old ally of compiler writers.
The notions of widening and narrowing were introduced by Cousot and Cousot in
one of the most cited papers in computer science~\cite{Cousot77}.
Different algorithms for range analysis have been subsequently proposed
by Stephenson {\em et al.}~\cite{Stephenson00}, Patterson~\cite{Patterson95}
and many other researchers.
Recently there have been many independent efforts to find exact, polynomial
time algorithms to solve constraints on the interval
lattice~\cite{Costan05,Gawlitza09,Lakhdar11,Su04,Su05}.
However, these works are still very theoretical, and have not yet been used to
analyze large programs.
Lakhdar {\em et al.}'s relational analysis~\cite{Lakhdar11}, for instance, takes
about 25 minutes to go over a program with almost 900 basic blocks.
We analyze programs of similar size in seconds.
We do not claim our approach is as precise as such algorithms, even though we
are able to exactly analyze 4/5 of the examples presented in~\cite{Lakhdar11}.
On the contrary, this paper presents a compromise between precision and speed
that scales to very large programs.

% Scalability: Astree and others.
There have been many practical approaches to abstract interpretation,
with special emphasis on range analysis~\cite{Blanchet03,Cousot09,Jung05}.
Cousot's group, for instance, has been able to globally analyze programs with
thousands of lines of code, albeit they use domain specific tools.
Astr\'{e}, for example, analyzes programs that do not contain recursive calls.
The work that is the closest to us is Oh {\em et al.}'s very recent abstract
interpretation framework~\cite{Oh12}.
Oh {\em et al.} discuss an implementation of range analysis on the interval
lattice that scales up to a program with 1,363K LoC (ghostscript-9.00), but
they do not provide results about its precision.
We could not find the benchmarks used in those experiments for a
direct comparison -- the distribution of ghostscript-9.00 available in the
LLVM test suite has 27K LoC.
On the other hand, we globally analyzed our largest benchmark, SPEC CPU 2006's
\texttt{gcc}, enabling function inlining, in less than 15secs.
Oh {\em et al.}'s implementation took orders of magnitude more time to go over
programs of similar size.
However, whereas they provide a framework to develop general sparse analyses, we
only solve range analysis on the interval lattice.

% PLDI 2012 paper

\section{Final Considerations}
\label{sec:con}

In this presentation we chose to omit correctness proofs for our algorithms.
For a proof that the widening and the narrowing operators give origin to
approximating sequences, we recommend Cousot and Cousot's work~\cite{Cousot77}.
For a proof that the e-SSA form transformation is semantics preserving, we
invite the reader to check a recent report of ours~\cite{Tavares11b}.
In this work we also show that the results that we obtain via the
sparse analysis are equivalent to the results provided by a dense analysis.
In other words, if the dense analysis tells us that variable $v$ is
associated with the interval range $[l, u]$ at program point $i$, and $v'$
is the new name of $v$, alive at $i$ in the e-SSA form program, then $v'$ is
associated with the interval $[l, u]$ along its entire live range.
Additionally, we have used a dynamic profiler, available in our webpage, to
empirically validate our results.

Our implementation of the range analysis algorithm described in this paper is
publicly available at \texttt{http://google.code/range-analysis/}.
This repository contains instructions about how to deploy and use our
implementation.
We provide a gallery of examples, including source codes,
CFGs and constraint graphs that we produce for meaningful programs at
\texttt{http://google.code/range-analysis/wiki/gallery}.

\bibliographystyle{plain}
\bibliography{references}

\end{document}