### 2012-04-09 doc/Manuscript/SBLP/paper.tex
Updated the files layout.

\documentclass{paper}
\usepackage{sbc-template}

\usepackage{proof}
\usepackage{mathpartir}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}

\def\qed{\unskip\kern 10pt{\unitlength1pt\linethickness{.4pt}\framebox(6,6){}}}

\newcommand{\fun}[1]{\mbox{\textbf{#1}}}
\newcommand{\lb}[1]{#1_{\downarrow}}
\newcommand{\ub}[1]{#1_{\uparrow}}
\newcommand{\varset}[1]{\mbox{$\cal{#1}$}}

\sloppy

\begin{document}

\title{Speed and Precision in Range Analysis}

\author{Victor Hugo Sperle Campos, Igor Rafael de Assis Costa,\\
Raphael Hernani Rodrigues and Fernando Magno Quint\~{a}o Pereira}

\address{UFMG -- 6627 Ant\^{o}nio Carlos Av, 31.270-010, Belo Horizonte, Brazil
\email{\{victorsc,igor.rafael,raphael.hernani,fernando\}@dcc.ufmg.br}
}

\maketitle

%mmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmm
\begin{abstract}
Range analysis is a compiler technique that determines statically the lower and
upper values that each integer variable from a target program may assume
during this program's execution.
This type of inference is very important, because it enables several compiler
optimizations, such as dead and redundant code elimination, bitwidth aware
register allocation, and detection of program vulnerabilities.
In this paper we describe an inter-procedural, context-sensitive range analysis
algorithm that we have implemented in the LLVM compiler.
During the effort to produce an industrial-quality implementation of our
algorithm, we had to face a constant tension between precision and speed.
The foremost goal of this paper is to discuss the many engineering choices
that, due to this tension, have shaped our implementation.
Given the breath of our evaluation, we believe that this paper
contains the most comprehensive empirical study of a range analysis
algorithm ever presented in the compiler related literature.
\end{abstract}

%mmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmm
\section{Introduction}
\label{sec:int}

% What range analysis is good for
Range analysis is a compiler technique whose objective is to determine
statically, for each program variable, limits for the minimum and maximum
values that this variable might assume during the program execution.
Range analysis is important because it enables many compiler optimizations.
Among these optimizations, the most well-known are dead and redundant
code elimination.
Examples of redundant code elimination include the removal of array bounds
checks~\cite{Bodik00,Logozzo08,Venet04} and overflow checks~\cite{Sol11}.
Additionally, range analysis is also used in bitwidth aware register
allocation~\cite{Barik06,Pereira08,Tallam03}, branch
prediction~\cite{Patterson95} and synthesis of hardware for specific
applications~\cite{Cong05,Lhairech10,Mahlke01,Stephenson00}.
Because of this importance, the programming language community has put much
effort in the design and implementation of efficient and precise range
analysis algorithms.

% The lack of a comprehensive empirical evaluation
However, the compiler related literature does not contain a comprehensive
evaluation of range analysis algorithms that scale up to entire programs.
Many works on this subject are limited to very small
programs~\cite{Mahlke01,Simon08,Stephenson00}, or, given their theoretic
perspective, have never been implemented in production
compilers~\cite{Costan05,Gawlitza09,Lakhdar11,Su04,Su05}.
There are implementations of range analysis that deal with very large
programs~\cite{Bertrane10,Cousot09,Logozzo08,Oh11}; nevertheless, because
these papers focus on applications of range analysis,
and not on its implementation, they do not provide a thorough discussion
about their engineering decisions.
A noticeable exception is the recent work of Oh {\em et al.}~\cite{Oh12},
which discusses a range analysis algorithm developed for C programs that can
handle very large benchmarks.
Oh {\em et al.} present an evaluation of the speed and memory consumption of
their implementation.
In this paper we claim to push this discussion considerably further.

% Summary of our experiments
We have implemented an industrial-quality range analysis algorithm in the
LLVM compiler~\cite{Lattner04}.
As we show in Section~\ref{sub:showdown}, our implementation, which is
publicly available, is able to analyze programs with over one million assembly
instructions in ten seconds.
And our implementation is not a straw-man: it produces very precise results.
We have compared the results produced by our implementation with the results
obtained via a dynamic profiler, which we have also implemented.
As we show in Section~\ref{sub:showdown}, when analyzing well-known numeric
benchmarks we are able to estimate tight ranges for almost half of all the
integer variables present in these programs.

% Design space exploration:
While designing and implementing our algorithm we had to face several important
engineering choices.
Many approaches that we have used in an attempt to increase the precision of
our implementation would result in runtime slowdowns.
Although we cannot determine the optimum spot in this design space, given the
vast number of possibilities, we discuss our most important implementation
decisions in Section~\ref{sec:design}.
Section~\ref{sub:sccs} shows how we could improve runtime and precision
substantially by processing data-flow information in the strongly connected
components that underly our constraint system.
Section~\ref{sub:program_rep} discuss the importance of choosing a suitable
intermediate representation when implementing a sparse data-flow framework.
Section~\ref{sub:whole} compares the intra-procedural and the inter-procedural
versions of our algorithm.
The role of context sensitiveness is discussed in Section~\ref{sub:context}.
Finally, Section~\ref{sub:widen} discusses the different widening strategies
that we have experimented with.

% The history of our implementation
This work concludes a two years long effort to produce a solid and scalable
implementation of range analysis.
Our first endeavor to implement such an algorithm was based on Su and
Wagner's constraint system, which can be solved exactly in polynomial
time~\cite{Su04,Su05}.
However, although we could use their formulation to handle a subset of C-like
constructs, their description of how to deal with loops was
not very explicit.
Thus, in order to solve loops we adopted Gawlitza
{\em et al.}'s~\cite{Gawlitza09} approach.
This technique uses the Bellman-Ford algorithm to detect increasing or
decreasing cycles in the constraint system, and then saturates these cycles
via a simple widening operator.
A detailed description of our implementation has been published by
Couto and Pereira~\cite{Couto11}.
Nevertheless, the inability to handle comparisons between variables, and the
cubic complexity of the Bellman-Ford method eventually led us to seek
alternative solutions to range analysis.
This quest reached a pinnacle in the present work, which we summarize in this
paper.

\section{Brief Description of our Range Analysis Algorithm}
\label{sec:desc}

% Define the lattice, the constraints and the valuation I.
\paragraph{The Interval Lattice.}
Following Gawlitza {\em et al.}'s notation, we shall be performing arithmetic
operations over the complete lattice
$\cal{Z} = \mathbb{Z} \cup \{-\infty, +\infty\}$, where the ordering is
naturally given by $-\infty < \ldots < -2 < -1 < 0 < 1 < 2 < \ldots +\infty$.
For any $x > -\infty$ we define:

\begin{tabular}{lcl}
$x + \infty = \infty, x \neq -\infty$ & \mbox{\hspace{0.1cm}} & $x - \infty = - \infty, x \neq +\infty$ \\
$x \times \infty = \infty$ if $x > 0$ & & $x \times \infty = -\infty$ if $x < 0$ \\
$0 \times \infty = 0$ & & $(-\infty) \times \infty = \ $ not defined $$ \\
\end{tabular}

From the lattice $\varset{Z}$ we define the product lattice
$\varset{Z}^2$, which is partially ordered by the subset relation
$\sqsubseteq$.
$\varset{Z}^2$ is defined as follows:
%
\begin{equation*}
\varset{Z}^2 = \emptyset \cup \{[z_1, z_2] | \ z_1,z_2 \in \varset{Z},
\ z_1 \leq z_2, \  -\infty < z_2 \}
\end{equation*}
%
The objective of range analysis is to determine a mapping
$I: \varset{V} \mapsto \varset{Z}^2$ from the set of integer program variables
$V$ to intervals, such that, for any variable $v \in V$, if
$I(v) = [l, u]$, the, during the execution of the target program, any
valued $i$ assigned to $v$ is such that $l \leq i \leq u$.

\paragraph{A Holistic View of our Range Analysis Algorithm.}
We perform range analysis in a number of steps.
First, we convert the program to a suitable intermediate representation that
makes it easier to extract constraints.
From these constraints, we build a dependence graph that allows us to do
range analysis sparsely.
Finally, we solve the constraints applying different fix-point iterators on
this dependence graph.
Figure~\ref{fig:algorithm} gives a global view of this algorithm.
Some of the steps in the algorithm are optional.
They improve the precision of the range analysis, at the expense of a longer
running time.
In Section~\ref{sec:design} we discuss in more details these tradeoffs.
The last phase, which we call the {\em micro algorithm}, happens per
strong component; however, if we opted for not building these components,
then it happens once for the entire constraint graph.
Nevertheless, the use of strongly connected components
is so essential for performance and precision, as we show in
Section~\ref{sub:sccs}, that it is considered optional only because we
can easily build our implementation without this module.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=\textwidth]{images/algorithm}
\end{center}
\caption{\label{fig:algorithm}
Our implementation of range analysis. Rounded boxes are optional steps.}
\end{figure}

We will illustrate the mandatory parts of the
algorithm via the example program in Figure~\ref{fig:overall_view}.
More details about each phase of the algorithm will be introduced in
Section~\ref{sec:design}, when we discuss our engineering decisions.
Figure~\ref{fig:overall_view}(a) shows an example program taken from the
partition function of the quicksort algorithm used by Bodik
{\em et al.}~\cite{Bodik00}.
We have removed the code that performs array manipulation from this program,
as it plays no role in our explanation.
Figure~\ref{fig:overall_view}(b) shows one possible way to represent this
program internally.
As we explain in Section~\ref{sub:program_rep}, a good program
representation allows us to find more precise results.
In this example we chose a program representation called
Extended Static Single Assignment form, which lets us to solve range
analysis via a path sensitive algorithm.
Figure~\ref{fig:overall_view}(c) shows the constraints that we extract from
the intermediate representation seen in part (b) of this figure.
From these constraints we build the {\em constraint graph} in
Figure~\ref{fig:overall_view}(d).
This graph is the main data-structure that we use to solve range analysis.
For each variable $v$ in the constraint system, the constraint graph has a node
$n_v$.
Similarly, for each constraint $v = f(\ldots, u, \ldots)$ in the constraint
system, the graph has an {\em operation node} $n_f$.
For each constraint $v = f(\ldots, u, \ldots)$ we add two edges to the
graph: $\overrightarrow{n_un_f}$ and $\overrightarrow{n_fn_v}$.
Some edges in the constraint graph are dashed.
These are called {\em control dependence edges}.
If a constraint $v = f(\ldots, \fun{ft}(u), \ldots)$ uses a {\em future}
bound from a variable $u$, then we add to the constraint graph a control
dependence edge $\overrightarrow{n_un_f}$.
The final solution to this instance of the range analysis problem is
given in Figure~\ref{fig:overall_view}(e).

\begin{figure}[t!]
\begin{center}
\includegraphics[width=\textwidth]{images/overall_view}
\end{center}
\caption{\label{fig:overall_view}
Range analysis by example.
(a) Input program.
(b) Internal compiler representation.
(c) Constraints of the range analysis problem.
(d) The constraint graph.
(e) The final solution.}
\end{figure}

\paragraph{The Micro Algorithm.}
We find the solution given in Figure~\ref{fig:overall_view}(e) in a
process that we call the micro algorithm.
This phase is further divided into three sub-steps:
(i) growth analysis;
(ii) future resolution and
(iii) narrowing analysis.

\noindent
\textbf{Growth analysis: }
The objective of growth analysis is to determine the growth behavior of
each program variable.
There are four possible behaviors:
(a) the variable is bound to a constant interval,
such as $k_0$ in Figure~\ref{fig:overall_view}(b).
(b) The variable is bound to a decreasing interval, i.e., an interval whose
lower bound decreases.
This is the case of $j_1$ in our example.
(c) The variable is bound to an increasing interval, i.e., its upper bound
increases.
This is the case of $i_1$ in the example.
(d) The variable is bound to an interval that expands in both directions.
The growth analysis uses an infinite lattice, i.e., $\varset{Z}^2$.
Thus, a careless implementation of an algorithm that infers growth patterns
might not terminate.
In order to ensure termination, we must reckon on a technique called
{\em widening}, first introduced by Cousot and Cousot as a key component
of abstract interpretation~\cite{Cousot77}.
There are many different widening strategies.
We discuss some of them in Section~\ref{sub:widen}.

\noindent
\textbf{Future resolution: }
In order to learn information from comparisons between variables, such as
\texttt{i < j} in Figure~\ref{fig:overall_view}(a), we bind some intervals
to {\em futures}.
Futures are symbolic limits, which will be replaced by actual numbers once
we finish the growth analysis.
The ranges found by the growth analysis tells us which variables have fixed
bounds, independent on the intersections in the constraint system.
Thus, we can use actual limits to replace intersections bounded by futures.
Figure~\ref{fig:fix_intersects} shows the rules to perform these substitutions.
In order to correctly replace a future $\fun{ft}(V)$ that limits a variable
$V'$, we need to have already applied the growth analysis onto $V$.
Had we considered only data dependence edges, then it would be possible
that $V'$ be analyzed before $V$.
However, because of control dependence edges, this case cannot happen.
The control dependence edges ensure that any topological ordering of the
constraint graph either places $N_v$ before $N_{v'}$, or places these nodes
in the same strongly connected component.
For instance, in Figure~\ref{fig:overall_view}(b), variables $j_1$ and $i_t$
are in the same SCC only because of the control dependence edges.

\begin{figure}[t!]
\begin{center}
\begin{eqnarray*}
\begin{array}{c}
\inferrule{Y = X \sqcap [l, \fun{ft}(V) + c] \\ \ub{I[V]} = u}
{Y = X \sqcap [l, u + c]} \mbox{\hspace{0.3cm}} u, c \in \mathbb{Z} \cup \{-\infty, +\infty\}
\\
\\
\inferrule{Y = X \sqcap [\fun{ft}(V) + c, u] \\ \lb{I[V]} = l}
{Y = X \sqcap [l + c, u]} \mbox{\hspace{0.3cm}} l, c \in \mathbb{Z} \cup \{-\infty, +\infty\}
\end{array}
\end{eqnarray*}
\end{center}
\caption{\label{fig:fix_intersects}Rules to replace futures by actual
bounds. $S$ is the interval bound to each variable after the widening
analysis.}
\end{figure}

\noindent
\textbf{Narrowing Analysis.}
The growth analysis associates very conservative bounds to each variable.
Thus, the last step of our algorithm consists in narrowing these intervals.
We accomplish this step via Cousot and Cousot's classic narrowing
operator~\cite[248]{Cousot77}, which we show in
Figure~\ref{fig:crop_analysis}.

\begin{figure}[t!]
\begin{center}
\begin{eqnarray*}
\begin{array}{c@{\hspace{0.9cm}}c}
\inferrule{\lb{I[Y]} = -\infty \\ \lb{e(Y)} > -\infty}
{I[Y] \leftarrow [\lb{e(Y)}, \ub{I[Y]}]}
&
\inferrule{\lb{I[Y]} > \lb{e(Y)}}
{I[Y] \leftarrow [\lb{e(Y)}, \ub{I[Y]}]}
\\
\\
\inferrule{\ub{I[Y]} = +\infty \\ \ub{e(Y)} < +\infty}
{I[Y] \leftarrow [\lb{I[Y]}, \ub{e(Y)}]}
&
\inferrule{\ub{I[Y]} < \ub{e(Y)}}
{I[Y] \leftarrow [\lb{I[Y]}, \ub{e(Y)}]}
\end{array}
\end{eqnarray*}
\end{center}
\caption{\label{fig:crop_analysis}Cousot and Cousot's narrowing operator.}
\end{figure}

\noindent
\textbf{Example:}
Continuing with our example, Figure~\ref{fig:ex_partition_grow_crop} shows
the application of our algorithm on the last strong component of
Figure~\ref{fig:overall_view}(d).
Upon meeting this SCC, we have already determined that the interval
$[0, 0]$ is bound to $i_0$ and that the interval $[100, 100]$ is bound to
$j_0$.
We are not guaranteed to find the least fix point of a constraint system.
However, in this example we did it.
We emphasize that finding this tight solution was only possible because of
the topological ordering of the constraint graph in
Figure~\ref{fig:overall_view}(d).
Had we applied the widening operator onto the whole graph, then we would
have found out that variable $j_0$ is bound to $[-\infty, +\infty]$,
because
(i) it receives its interval directly from variable $k_t$, which is upper
bounded by $+\infty$, and
(ii) it is part of a negative cycle.
On the other hand, by only analyzing $j$'s SCC after we have
analyzed $k$'s, $k$ only contribute the constant range $[0, 99]$ to $j_0$.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=\textwidth]{images/ex_partition_grow_crop}
\end{center}
\caption{\label{fig:ex_partition_grow_crop}
Four snapshots of the last SCC of Figure~\ref{fig:overall_view}(d).
(a) After removing control dependence edges.
(b) After running the growth analysis.
(c) After fixing the intersections bound to futures.
(d) After running the narrowing analysis.}
\end{figure}

\subsection{Range Analysis Showdown}
\label{sub:showdown}

The objective of this section is to show, via experimental numbers, that our
implementation of range analysis is fast, economic and effective.
We have used it to analyze a test suite with 2.72 million lines of C code,
which includes, in addition to all the benchmarks distributed with LLVM,
the programs in the SPEC CPU 2006 collection.

\paragraph{Time and Memory Complexity}

Figure~\ref{fig:TimeCorr} provides a visual comparison between the time to
run our algorithm and the size of the input programs.
We show data for the 100 largest benchmarks in our test suite, in number
of variable nodes in the constraint graph.
We perform function inlining before running our analysis, to increase program
sizes.
Each point in the X line corresponds to a benchmark.
We analyze the smallest benchmark in this set, \texttt{Prolangs-C/deriv2}, which
has 1,131 variable nodes in the constraint graph, in 20ms.
We take 15,91secs to analyze our largest benchmark, \texttt{403.gcc}, which,
after function inlining, has 1,266,273 assembly instructions, and a
constraint graph with 679,652 variable nodes.
For this data set, the coefficient of determination $(R^2)$ is 0.967, which
provides very strong evidence about the linear asymptotic complexity of our
implementation.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=0.9\textwidth]{images/TimeCorr}
\end{center}
\caption{\label{fig:TimeCorr}
Correlation between program size (number of var nodes in constraint
graphs after inlining) and analysis runtime (ms).
Coefficient of determination = 0.967.
}
\end{figure}

\noindent
\textbf{Memory complexity.}
The experiments also reveal that the memory consumption of our implementation
is linear with the program size.
Figure~\ref{fig:MemCorr} plots these two quantities together.
The linear correlation, in this case, is even stronger than that found in
Figure~\ref{fig:TimeCorr}, which compares runtime and program size: the
coefficient of determination is 0.9947.
The figure only shows our 100 largest benchmarks.
Again, SPEC \texttt{403.gcc} is the heaviest benchmarks, requiring
265,588KB to run.
Memory includes stack, heap and the executable program code.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=0.9\textwidth]{images/MemCorr}
\end{center}
\caption{\label{fig:MemCorr}
Comparison between program size (number of var nodes in constraint
graphs after inlining) and memory consumption (KB).
Coefficient of determination = 0.9947.
}
\end{figure}

\paragraph{Precision}

Our implementation of range analysis is remarkably precise, considering its
runtime.
Lakhdar {\em et al.}'s relational analysis~\cite{Lakhdar11}, for instance, takes
about 25 minutes to go over a program with almost 900 basic blocks.
We analyze programs of similar size less than one second.
We do not claim our approach is as precise as such algorithms, even though we
are able to find exact bounds to 4/5 of the examples presented
in~\cite{Lakhdar11}.
On the contrary, this paper presents a compromise between precision and speed
that scales to very large programs.
Nevertheless, our results are far from being trivial.
We have implemented a dynamic profiler that measures, for each variable,
its upper and lower limits, given an execution of the target program.
Figure~\ref{fig:precision} compares our results with those measured
dynamically for the Stanford benchmark suite, which is publicly
available~\footnote{\texttt{http://classes.engineering.wustl.edu/cse465/docs/BCCExamples/stanford.c}}.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=\textwidth]{images/precUpperBound}
\includegraphics[width=\textwidth]{images/precLowerBound}
\end{center}
\caption{\label{fig:precision}
(Upper) Comparison between static range analysis and dynamic profiler for
upper bounds.
(Lower) Comparison between static range analysis and dynamic profiler for
lower bounds. The numbers above the benchmark names give the number of
variables in each program.}
\end{figure}

We have classified the bounds estimated by the static analysis into four
categories.
The first category, which we call $1$, contains those bounds that are tight:
during the execution of the program, the variable has been assigned an upper,
or lower limit, that equals the limit inferred statically.
The second category, which we call $n$, contains the estimated bounds that are
within twice the value inferred statically.
For instance, if the range analysis estimates that a variable $v$ is in the
range $[0, 100]$, and during the execution the dynamic profiler finds that
its maximum value is $51$, then $v$ falls into this category.
The third category, $n^2$, contains variables whose actual value is within
a quadratic factor from the estimated value.
In our example, $v$'s upper bound would have to be at most $10$ for it to
be in this category.
Finally, the fourth category contains variables whose estimated value lays
outside a quadratic factor of the actual value.
We call this category {\em imprecise}, and it contains mostly the limits that
our static analysis has marked as either $+\infty$ or $-\infty$.

As we see in Figure~\ref{fig:precision}, 54.11\% of the lower limits that
we have estimated statically are exact.
Similarly, 51.99\% of our upper bounds are also tight.
The figure also shows that, on average, 37.39\% of our lower limits are
imprecise, and 35.40\% of our upper limits are imprecise.
This result is on pair with those obtained by more costly analysis, such as
Stephenson {\em et al.}'s~\cite{Stephenson00}.
However, whereas that approach have not been used with programs larger than
the Stanford benchmark suite, we, as shown before, have been able to
deal with remarkably large programs.

%mmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmm
\section{Design Space}
\label{sec:design}

As we see from a cursory glance at Figure~\ref{fig:algorithm}, our range
analysis algorithm has many optional modules.
These modules provide the user with the possibility to choose between
more precise results, or a faster analysis.
Given the number of options, the design space of a range analysis algorithm
is vast.
In this section we try to cover some of the most important tradeoffs.
Figure~\ref{fig:space} plots, for the integer programs in the SPEC CPU 2006
benchmark suite, precision versus speed for different configurations of
our implementation.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=\textwidth]{images/space}
\end{center}
\caption{\label{fig:space}
Design space exploration: precision versus speed for different configurations
of our algorithm.}
\end{figure}

\subsection{Strongly Connected Components}
\label{sub:sccs}

\subsection{The choice of a program representation}
\label{sub:program_rep}

\subsection{Intra versus Inter-procedural Analysis}
\label{sub:whole}

\subsection{Context Sensitive versus Context Insensitive Analysis}
\label{sub:context}

\subsection{Choosing a Widening Strategy}
\label{sub:widen}

%mmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmm
\section{Conclusion}
\label{sec:con}

\bibliographystyle{plain}
\bibliography{../references}

\end{document}